Making Sequential Programs Deterministic

### 2019-04-06 related.tex
final edits

The state of the art for detecting nondeterministic tests is NonDex~\cite{nondex}.  It is a highly effective tool.
% https://github.com/TestingResearchIllinois/NonDex/wiki/Supported-APIs
NonDex uses a hand-crafted list of 47 methods (25 unique method names)
in 13 classes as potential sources of flakiness. 
For each of the identified methods, the authors built models that
return different results when called consecutively. A modified JVM then
runs a given test multiple times and reports the test as being flaky if it observes
diverging test output. While this approach produces precise results, it requires manual inspection
and considerable debugging effort to locate and fix the source of flakiness. \TheDeterminismChecker, in
contrast, reports the cause of nondeterminism at compile time requiring
little to no debugging effort.  However, \theDeterminismChecker requires
much more upfront effort, and it produce false positive warnings.
NonDex's approach of identifying and modeling methods with nondeterministic
specifications is analogous to our library specifications. So far, we have annotated
928 methods across 41 classes in the JDK, in addition to other libraries.
%\todo{Check whether our JDK is missing anything they support - No}
%\todo{Did the NonDex authors contact the maintainers of the tools in which
%  they found nondeterminism - No!}
%\todo{The NonDex authors didn't contact the maintainers of tools.
%From the paper: "NonDex detected 57 previously unknown
%flaky tests in open-source projects and three flaky tests
%that have been already fixed by the open-source software
%developers."}

@@

Several techniques have been proposed to test whether a
deterministic implementation conforms to its nondeterministic finite state machine~\cite{Petrenko1996,Petrenko:1993:NSM:648128.761244,Savor:1997:639710,Hierons:2004:TCD:1040993.1040998}.
\cite{Cook:2013:RNP:2491956.2491969} presents an approach that can automatically verify properties
in branching time temporal logic systems that are inherently
nondeterministic.
Bocchino et
al.~\cite{Bocchino:2009:TES:1640089.1640097,Bocchino:2011:SND:1926385.1926447}
provide a type-and-effect system that provides compile-time determinism
guarantees for parallel programs, with a focus on barrier removal,
reasoning about interference and thread interleavings.
We believe our work to be the first verification approach
aimed at ensuring determinism in sequential programs.

>>>>

The state of the art for detecting nondeterministic tests is NonDex~\cite{nondex}.  It is a highly effective tool.
% https://github.com/TestingResearchIllinois/NonDex/wiki/Supported-APIs
NonDex uses a hand-crafted list of 47 methods (25 unique method names)
in 13 classes as potential sources of flakiness. 
For each of the identified methods, the authors built models that
return different results when called consecutively. A modified JVM then
runs a given test multiple times and reports the test as being flaky if it observes
diverging test output. While this approach produces precise results, it requires manual inspection
and considerable debugging effort to locate and fix the source of flakiness. \TheDeterminismChecker, in
contrast, reports the cause of nondeterminism at compile time requiring
little to no debugging effort.  However, \theDeterminismChecker requires
much more upfront effort, and it produces false positive warnings.
NonDex's approach of identifying and modeling methods with nondeterministic
specifications is analogous to our library specifications. So far, we have annotated
928 methods across 41 classes in the JDK, in addition to other libraries.
%\todo{Check whether our JDK is missing anything they support - No}
%\todo{Did the NonDex authors contact the maintainers of the tools in which
%  they found nondeterminism - No!}
%\todo{The NonDex authors didn't contact the maintainers of tools.
%From the paper: "NonDex detected 57 previously unknown
%flaky tests in open-source projects and three flaky tests
%that have been already fixed by the open-source software
%developers."}

@@

Several techniques have been proposed to test whether a
deterministic implementation conforms to its nondeterministic finite state machine~\cite{Petrenko1996,Petrenko:1993:NSM:648128.761244,Savor:1997:639710,Hierons:2004:TCD:1040993.1040998}.
\cite{Cook:2013:RNP:2491956.2491969} presents an approach that can automatically verify properties
in branching time temporal logic systems that are inherently
nondeterministic.
Bocchino et
al.~\cite{Bocchino:2009:TES:1640089.1640097,Bocchino:2011:SND:1926385.1926447}
present a type-and-effect system that provides compile-time determinism
guarantees for parallel programs, with a focus on barrier removal,
reasoning about interference and thread interleavings.
We believe our work to be the first verification approach
aimed at ensuring determinism in sequential programs.
